\bf Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-04-09 paper.qmd
fixed space after pD

Non-linear dimension reduction (NLDR) is popular for making a convenient low-dimensional (\kD) representation of high-dimensional (\pD) data. Recently developed methods include t-distributed stochastic neighbor embedding (tSNE) [@laurens2008], uniform manifold approximation and projection (UMAP) [@leland2018], potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm [@moon2019], large-scale dimensionality reduction Using triplets (TriMAP) [@amid2022], and pairwise controlled manifold approximation (PaCMAP) [@yingfan2021]. However, the representation generated can vary dramatically from method to method, and with different choices of parameters or random seeds made using the same method (@fig-NLDR-variety). The dilemma for the analyst is then, **which representation to use**. The choice might result in different procedures used in the downstream analysis, or different inferential conclusions. The research described here provides new visual tools to aid with this decision.

@@

<!-- - Connection between NLDR and MDS-->
Historically, \kD   representations of \pD   data have been computed using multidimensional scaling (MDS) [@borg2005], which includes principal components analysis (PCA) [@jolliffe2011] as a special case.  The \kD   representation can be considered to be a layout of points in \kD   produced by an embedding procedure that maps the data from \pD. In MDS, the \kD   layout is constructed by minimizing a stress function that differences distances between points in \pD   with potential distances between points in \kD. Various formulations of the stress function result in non-metric scaling [@saeed2018] and isomap [@silva2002]. Challenges in working with high-dimensional data, including visualization, are outlined in @johnstone2009.

Many new methods for NLDR have emerged in recent years, all designed to better capture specific structures potentially existing in \pD. Here we focus on five currently popular techniques, tSNE, UMAP, PHATE, TriMAP and PaCMAP. tNSE and UMAP can be considered to produce the \kD   minimizing the divergence between two distributions, where the distributions are modeling the inter-point distances. PHATE, TriMAP and PaCMAP are examples of diffusion processes [@coifman2005] spreading to capture geometric shapes, that include both global and local structure.

@@

The alternative approach to visualizing the high-dimensional data is to use linear projections. PCA is the classical approach, resulting in a set of new variables which are linear combinations of the original variables. Tours, defined by @lee2021, broaden the scope by providing movies of linear projections, that provide views the data from all directions. @lee2021 provides an review of the main developments in tours. There are many tour algorithms implemented, with many available in the R package `tourr` [@wickham2011], and versions enabling better interactivity in `langevitour` [@harisson2024] and `detourr` [@hart2022]. Linear projections are a safe way to view high-dimensional data, because they do not warp the space, so they are more faithful representations of the structure. 
However, linear projections can be cluttered, and global patterns can obscure local structure. The simple activity of projecting data from \pD   suffers from piling [@laa2022], where data concentrates in the center of projections. NLDR is designed to escape these issues, to exaggerate structure so that it can be observed. But as a result NLDR can hallucinate wildly, to suggest patterns that are not actually present in the data.

The solution is to use the tour to examine how the NLDR is warping the space. This approach follows what @wickham2015 describes as *model-in-the-data-space*. The fitted model should be overlaid on the data, to examine the fit relative the spread of the observations. While this is straightforward, and commonly done when data is 2D, it is also possible in \pD, for many models, when a tour is used.

@wickham2015 provides several examples of models overlaid on the data in \pD. In hierarchical clustering, a representation of the dendrogrom using points and lines can be constructed by augmenting the data with points marking merging of clusters. Showing the movie of linear projections reveals shows how the algorithm sequentially fitted the cluster model to the data. For linear discriminant analysis or model-based clustering the model can be indicated by $(p-1)\text{-}D$ ellipses. It is possible to see whether the elliptical shapes appropriately matches the variance of the relevant clusters, and to compare and contrast different fits. For PCA, one can display the \kD plane of the reduced dimension using wireframes of transformed cubes. Using a wireframe is the approach we take here, to represent the NLDR model in \pD.

@@

At first glance, thinking of NLDR as a modeling technique might seem strange. It is a simplified representation or abstraction of a system, process, or phenomenon in the real world. The \pD   observations are the realization of the phenomenon, and the \kD   NLDR layout is the simplified representation. From a statistical perspective we can consider the distances between points in the \kD   layout to be variance that the model explains, and the (relative) difference with their distances in \pD   is the error, or unexplained variance. We can also imagine that the positioning of points in 2D represent the fitted values, that will have some prescribed position in \pD   that can be compared with their observed values. This is the conceptual framework underlying the more formal versions of factor analysis [@cfa69] and multidimensional scaling (MDS) [@borg2005]. (Note that, for this thinking the full \pD   data needs to be available, not just the interpoint distances.)
<!--### Notation -->

@@

With our goal being to make a representation of this 2D layout that can be lifted into high-dimensional space, the layout needs to be augmented to include neighbour information. A simple approach would be to triangulate the points and add edges. A more stable approach is to first hexagonally bin the data, and connect the bin centroids. This process serves reduce some noise in the resulting surface shown in \pD.

Define a hexagonal grid across the \kD space, using hexagons with fixed height ($h_b$), width ($w_b$). The starting coordinates $(x_s, y_s)$ mark the lower left of the grid. <!--We deliberately separate out the creation of the hexagonal grid from the mapping of points on the grid.-->
Each of the \kD points will belong to a hexagonal bin. That is, for each $y \in \mathbfit{Y}$, we can (uniquely) identify the hexagon that the point belongs to. Let $h : \mathbb{R}^k \rightarrow \mathbb{R}^k$ be the function that maps a point in \kD space to it's hexagonal centroid ($C^{(k)}$). <!--It follows that $h \circ g : \mathbb{R}^p \rightarrow \mathbb{R}^k$ maps each point in $\mathbfit{X}$ to a \kD centroid.-->

@@

### Displaying the model in \pD

The final step of the algorithm is _lifting_ the \kD model back to \pD, completing the *model-in-the-data-space* picture. For this, define the set $H^k$ to be the set of points that belong to centroid $k$. That is, $H^k = \{ y \; | \; g(y) = (h_x^k, h_y^k)\}$. We use the \pD Euclidean mean of the points in $H^k$ to map the centroid $(h_x^k, h_y^k)$ to a point in \pD. Let the $i$-th component of the \pD mean be

@@

with corresponding \pD vector $\vec{f_i}$. Then, $f(h\circ g)(x)$ maps a \pD point to the \pD model estimate, completing the model cycle.

>>>>

Non-linear dimension reduction (NLDR) is popular for making a convenient low-dimensional (\kD{}) representation of high-dimensional (\pD{}) data. Recently developed methods include t-distributed stochastic neighbor embedding (tSNE) [@laurens2008], uniform manifold approximation and projection (UMAP) [@leland2018], potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm [@moon2019], large-scale dimensionality reduction Using triplets (TriMAP) [@amid2022], and pairwise controlled manifold approximation (PaCMAP) [@yingfan2021]. However, the representation generated can vary dramatically from method to method, and with different choices of parameters or random seeds made using the same method (@fig-NLDR-variety). The dilemma for the analyst is then, **which representation to use**. The choice might result in different procedures used in the downstream analysis, or different inferential conclusions. The research described here provides new visual tools to aid with this decision.

@@

<!-- - Connection between NLDR and MDS-->
Historically, \kD{}   representations of \pD{}   data have been computed using multidimensional scaling (MDS) [@borg2005], which includes principal components analysis (PCA) [@jolliffe2011] as a special case.  The \kD{}   representation can be considered to be a layout of points in \kD{}   produced by an embedding procedure that maps the data from \pD{}. In MDS, the \kD{}   layout is constructed by minimizing a stress function that differences distances between points in \pD{}   with potential distances between points in \kD{}. Various formulations of the stress function result in non-metric scaling [@saeed2018] and isomap [@silva2002]. Challenges in working with high-dimensional data, including visualization, are outlined in @johnstone2009.

Many new methods for NLDR have emerged in recent years, all designed to better capture specific structures potentially existing in \pD{}. Here we focus on five currently popular techniques, tSNE, UMAP, PHATE, TriMAP and PaCMAP. tNSE and UMAP can be considered to produce the \kD{}   minimizing the divergence between two distributions, where the distributions are modeling the inter-point distances. PHATE, TriMAP and PaCMAP are examples of diffusion processes [@coifman2005] spreading to capture geometric shapes, that include both global and local structure.

@@

The alternative approach to visualizing the high-dimensional data is to use linear projections. PCA is the classical approach, resulting in a set of new variables which are linear combinations of the original variables. Tours, defined by @lee2021, broaden the scope by providing movies of linear projections, that provide views the data from all directions. @lee2021 provides an review of the main developments in tours. There are many tour algorithms implemented, with many available in the R package `tourr` [@wickham2011], and versions enabling better interactivity in `langevitour` [@harisson2024] and `detourr` [@hart2022]. Linear projections are a safe way to view high-dimensional data, because they do not warp the space, so they are more faithful representations of the structure. 
However, linear projections can be cluttered, and global patterns can obscure local structure. The simple activity of projecting data from \pD{}   suffers from piling [@laa2022], where data concentrates in the center of projections. NLDR is designed to escape these issues, to exaggerate structure so that it can be observed. But as a result NLDR can hallucinate wildly, to suggest patterns that are not actually present in the data.

The solution is to use the tour to examine how the NLDR is warping the space. This approach follows what @wickham2015 describes as *model-in-the-data-space*. The fitted model should be overlaid on the data, to examine the fit relative the spread of the observations. While this is straightforward, and commonly done when data is 2D, it is also possible in \pD{}, for many models, when a tour is used.

@wickham2015 provides several examples of models overlaid on the data in \pD{}. In hierarchical clustering, a representation of the dendrogrom using points and lines can be constructed by augmenting the data with points marking merging of clusters. Showing the movie of linear projections reveals shows how the algorithm sequentially fitted the cluster model to the data. For linear discriminant analysis or model-based clustering the model can be indicated by $(p-1)\text{-}D$ ellipses. It is possible to see whether the elliptical shapes appropriately matches the variance of the relevant clusters, and to compare and contrast different fits. For PCA, one can display the \kD{} plane of the reduced dimension using wireframes of transformed cubes. Using a wireframe is the approach we take here, to represent the NLDR model in \pD{}.

@@

At first glance, thinking of NLDR as a modeling technique might seem strange. It is a simplified representation or abstraction of a system, process, or phenomenon in the real world. The \pD{}   observations are the realization of the phenomenon, and the \kD{}   NLDR layout is the simplified representation. From a statistical perspective we can consider the distances between points in the \kD{}   layout to be variance that the model explains, and the (relative) difference with their distances in \pD{}   is the error, or unexplained variance. We can also imagine that the positioning of points in 2D represent the fitted values, that will have some prescribed position in \pD{}   that can be compared with their observed values. This is the conceptual framework underlying the more formal versions of factor analysis [@cfa69] and multidimensional scaling (MDS) [@borg2005]. (Note that, for this thinking the full \pD{}   data needs to be available, not just the interpoint distances.)
<!--### Notation -->

@@

With our goal being to make a representation of this 2D layout that can be lifted into high-dimensional space, the layout needs to be augmented to include neighbour information. A simple approach would be to triangulate the points and add edges. A more stable approach is to first hexagonally bin the data, and connect the bin centroids. This process serves reduce some noise in the resulting surface shown in \pD{}.

Define a hexagonal grid across the \kD{} space, using hexagons with fixed height ($h_b$), width ($w_b$). The starting coordinates $(x_s, y_s)$ mark the lower left of the grid. <!--We deliberately separate out the creation of the hexagonal grid from the mapping of points on the grid.-->
Each of the \kD{} points will belong to a hexagonal bin. That is, for each $y \in \mathbfit{Y}$, we can (uniquely) identify the hexagon that the point belongs to. Let $h : \mathbb{R}^k \rightarrow \mathbb{R}^k$ be the function that maps a point in \kD{} space to it's hexagonal centroid ($C^{(k)}$). <!--It follows that $h \circ g : \mathbb{R}^p \rightarrow \mathbb{R}^k$ maps each point in $\mathbfit{X}$ to a \kD{} centroid.-->

@@

### Displaying the model in \pD{}

The final step of the algorithm is _lifting_ the \kD{} model back to \pD{}, completing the *model-in-the-data-space* picture. For this, define the set $H^k$ to be the set of points that belong to centroid $k$. That is, $H^k = \{ y \; | \; g(y) = (h_x^k, h_y^k)\}$. We use the \pD{} Euclidean mean of the points in $H^k$ to map the centroid $(h_x^k, h_y^k)$ to a point in \pD{}. Let the $i$-th component of the \pD{} mean be

@@

with corresponding \pD{} vector $\vec{f_i}$. Then, $f(h\circ g)(x)$ maps a \pD{} point to the \pD{} model estimate, completing the model cycle.

### 2024-04-09 paper.tex
fixed space after pD

Non-linear dimension reduction (NLDR) is popular for making a convenient
low-dimensional (\kD) representation of high-dimensional (\pD) data.
Recently developed methods include t-distributed stochastic neighbor
embedding (tSNE) \citep{laurens2008}, uniform manifold approximation and
projection (UMAP) \citep{leland2018}, potential of heat-diffusion for
affinity-based trajectory embedding (PHATE) algorithm \citep{moon2019},
large-scale dimensionality reduction Using triplets (TriMAP)
\citep{amid2022}, and pairwise controlled manifold approximation
(PaCMAP) \citep{yingfan2021}. However, the representation generated can
vary dramatically from method to method, and with different choices of
parameters or random seeds made using the same method
(Figure~\ref{fig-NLDR-variety}). The dilemma for the analyst is then,
\textbf{which representation to use}. The choice might result in
different procedures used in the downstream analysis, or different
inferential conclusions. The research described here provides new visual
tools to aid with this decision.

@@

Historically, \kD   representations of \pD   data have been computed
using multidimensional scaling (MDS) \citep{borg2005}, which includes
principal components analysis (PCA) \citep{jolliffe2011} as a special
case. The \kD   representation can be considered to be a layout of
points in \kD   produced by an embedding procedure that maps the data
from \pD. In MDS, the \kD   layout is constructed by minimizing a stress
function that differences distances between points in \pD   with
potential distances between points in \kD. Various formulations of the
stress function result in non-metric scaling \citep{saeed2018} and
isomap \citep{silva2002}. Challenges in working with high-dimensional
data, including visualization, are outlined in \citet{johnstone2009}.

Many new methods for NLDR have emerged in recent years, all designed to
better capture specific structures potentially existing in \pD. Here we
focus on five currently popular techniques, tSNE, UMAP, PHATE, TriMAP
and PaCMAP. tNSE and UMAP can be considered to produce the
\kD   minimizing the divergence between two distributions, where the
distributions are modeling the inter-point distances. PHATE, TriMAP and
PaCMAP are examples of diffusion processes \citep{coifman2005} spreading
to capture geometric shapes, that include both global and local
structure.

@@

The alternative approach to visualizing the high-dimensional data is to
use linear projections. PCA is the classical approach, resulting in a
set of new variables which are linear combinations of the original
variables. Tours, defined by \citet{lee2021}, broaden the scope by
providing movies of linear projections, that provide views the data from
all directions. \citet{lee2021} provides an review of the main
developments in tours. There are many tour algorithms implemented, with
many available in the R package \texttt{tourr} \citep{wickham2011}, and
versions enabling better interactivity in \texttt{langevitour}
\citep{harisson2024} and \texttt{detourr} \citep{hart2022}. Linear
projections are a safe way to view high-dimensional data, because they
do not warp the space, so they are more faithful representations of the
structure. However, linear projections can be cluttered, and global
patterns can obscure local structure. The simple activity of projecting
data from \pD   suffers from piling \citep{laa2022}, where data
concentrates in the center of projections. NLDR is designed to escape
these issues, to exaggerate structure so that it can be observed. But as
a result NLDR can hallucinate wildly, to suggest patterns that are not
actually present in the data.

The solution is to use the tour to examine how the NLDR is warping the
space. This approach follows what \citet{wickham2015} describes as
\emph{model-in-the-data-space}. The fitted model should be overlaid on
the data, to examine the fit relative the spread of the observations.
While this is straightforward, and commonly done when data is 2D, it is
also possible in \pD, for many models, when a tour is used.

\citet{wickham2015} provides several examples of models overlaid on the
data in \pD. In hierarchical clustering, a representation of the
dendrogrom using points and lines can be constructed by augmenting the
data with points marking merging of clusters. Showing the movie of
linear projections reveals shows how the algorithm sequentially fitted
the cluster model to the data. For linear discriminant analysis or
model-based clustering the model can be indicated by \((p-1)\text{-}D\)
ellipses. It is possible to see whether the elliptical shapes
appropriately matches the variance of the relevant clusters, and to
compare and contrast different fits. For PCA, one can display the
\kD plane of the reduced dimension using wireframes of transformed
cubes. Using a wireframe is the approach we take here, to represent the
NLDR model in \pD.

@@

At first glance, thinking of NLDR as a modeling technique might seem
strange. It is a simplified representation or abstraction of a system,
process, or phenomenon in the real world. The \pD   observations are the
realization of the phenomenon, and the \kD   NLDR layout is the
simplified representation. From a statistical perspective we can
consider the distances between points in the \kD   layout to be variance
that the model explains, and the (relative) difference with their
distances in \pD   is the error, or unexplained variance. We can also
imagine that the positioning of points in 2D represent the fitted
values, that will have some prescribed position in \pD   that can be
compared with their observed values. This is the conceptual framework
underlying the more formal versions of factor analysis \citep{cfa69} and
multidimensional scaling (MDS) \citep{borg2005}. (Note that, for this
thinking the full \pD   data needs to be available, not just the
interpoint distances.)

@@

With our goal being to make a representation of this 2D layout that can
be lifted into high-dimensional space, the layout needs to be augmented
to include neighbour information. A simple approach would be to
triangulate the points and add edges. A more stable approach is to first
hexagonally bin the data, and connect the bin centroids. This process
serves reduce some noise in the resulting surface shown in \pD.

Define a hexagonal grid across the \kD space, using hexagons with fixed
height (\(h_b\)), width (\(w_b\)). The starting coordinates
\((x_s, y_s)\) mark the lower left of the grid. Each of the \kD points
will belong to a hexagonal bin. That is, for each
\(y \in \mathbfit{Y}\), we can (uniquely) identify the hexagon that the
point belongs to. Let \(h : \mathbb{R}^k \rightarrow \mathbb{R}^k\) be
the function that maps a point in \kD space to it's hexagonal centroid
(\(C^{(k)}\)).

@@

\subsection{\texorpdfstring{Displaying the model in
\pD}{Displaying the model in }}\label{displaying-the-model-in}

The final step of the algorithm is \emph{lifting} the \kD model back to
\pD, completing the \emph{model-in-the-data-space} picture. For this,
define the set \(H^k\) to be the set of points that belong to centroid
\(k\). That is, \(H^k = \{ y \; | \; g(y) = (h_x^k, h_y^k)\}\). We use
the \pD Euclidean mean of the points in \(H^k\) to map the centroid
\((h_x^k, h_y^k)\) to a point in \pD. Let the \(i\)-th component of the
\pD mean be

@@

with corresponding \pD vector \(\vec{f_i}\). Then, \(f(h\circ g)(x)\)
maps a \pD point to the \pD model estimate, completing the model cycle.

>>>>

Non-linear dimension reduction (NLDR) is popular for making a convenient
low-dimensional (\kD{}) representation of high-dimensional (\pD{}) data.
Recently developed methods include t-distributed stochastic neighbor
embedding (tSNE) \citep{laurens2008}, uniform manifold approximation and
projection (UMAP) \citep{leland2018}, potential of heat-diffusion for
affinity-based trajectory embedding (PHATE) algorithm \citep{moon2019},
large-scale dimensionality reduction Using triplets (TriMAP)
\citep{amid2022}, and pairwise controlled manifold approximation
(PaCMAP) \citep{yingfan2021}. However, the representation generated can
vary dramatically from method to method, and with different choices of
parameters or random seeds made using the same method
(Figure~\ref{fig-NLDR-variety}). The dilemma for the analyst is then,
\textbf{which representation to use}. The choice might result in
different procedures used in the downstream analysis, or different
inferential conclusions. The research described here provides new visual
tools to aid with this decision.

@@

Historically, \kD{} representations of \pD{} data have been computed
using multidimensional scaling (MDS) \citep{borg2005}, which includes
principal components analysis (PCA) \citep{jolliffe2011} as a special
case. The \kD{} representation can be considered to be a layout of
points in \kD{} produced by an embedding procedure that maps the data
from \pD{}. In MDS, the \kD{} layout is constructed by minimizing a
stress function that differences distances between points in \pD{} with
potential distances between points in \kD{}. Various formulations of the
stress function result in non-metric scaling \citep{saeed2018} and
isomap \citep{silva2002}. Challenges in working with high-dimensional
data, including visualization, are outlined in \citet{johnstone2009}.

Many new methods for NLDR have emerged in recent years, all designed to
better capture specific structures potentially existing in \pD{}. Here
we focus on five currently popular techniques, tSNE, UMAP, PHATE, TriMAP
and PaCMAP. tNSE and UMAP can be considered to produce the \kD{}
minimizing the divergence between two distributions, where the
distributions are modeling the inter-point distances. PHATE, TriMAP and
PaCMAP are examples of diffusion processes \citep{coifman2005} spreading
to capture geometric shapes, that include both global and local
structure.

@@

The alternative approach to visualizing the high-dimensional data is to
use linear projections. PCA is the classical approach, resulting in a
set of new variables which are linear combinations of the original
variables. Tours, defined by \citet{lee2021}, broaden the scope by
providing movies of linear projections, that provide views the data from
all directions. \citet{lee2021} provides an review of the main
developments in tours. There are many tour algorithms implemented, with
many available in the R package \texttt{tourr} \citep{wickham2011}, and
versions enabling better interactivity in \texttt{langevitour}
\citep{harisson2024} and \texttt{detourr} \citep{hart2022}. Linear
projections are a safe way to view high-dimensional data, because they
do not warp the space, so they are more faithful representations of the
structure. However, linear projections can be cluttered, and global
patterns can obscure local structure. The simple activity of projecting
data from \pD{} suffers from piling \citep{laa2022}, where data
concentrates in the center of projections. NLDR is designed to escape
these issues, to exaggerate structure so that it can be observed. But as
a result NLDR can hallucinate wildly, to suggest patterns that are not
actually present in the data.

The solution is to use the tour to examine how the NLDR is warping the
space. This approach follows what \citet{wickham2015} describes as
\emph{model-in-the-data-space}. The fitted model should be overlaid on
the data, to examine the fit relative the spread of the observations.
While this is straightforward, and commonly done when data is 2D, it is
also possible in \pD{}, for many models, when a tour is used.

\citet{wickham2015} provides several examples of models overlaid on the
data in \pD{}. In hierarchical clustering, a representation of the
dendrogrom using points and lines can be constructed by augmenting the
data with points marking merging of clusters. Showing the movie of
linear projections reveals shows how the algorithm sequentially fitted
the cluster model to the data. For linear discriminant analysis or
model-based clustering the model can be indicated by \((p-1)\text{-}D\)
ellipses. It is possible to see whether the elliptical shapes
appropriately matches the variance of the relevant clusters, and to
compare and contrast different fits. For PCA, one can display the \kD{}
plane of the reduced dimension using wireframes of transformed cubes.
Using a wireframe is the approach we take here, to represent the NLDR
model in \pD{}.

@@

At first glance, thinking of NLDR as a modeling technique might seem
strange. It is a simplified representation or abstraction of a system,
process, or phenomenon in the real world. The \pD{} observations are the
realization of the phenomenon, and the \kD{} NLDR layout is the
simplified representation. From a statistical perspective we can
consider the distances between points in the \kD{} layout to be variance
that the model explains, and the (relative) difference with their
distances in \pD{} is the error, or unexplained variance. We can also
imagine that the positioning of points in 2D represent the fitted
values, that will have some prescribed position in \pD{} that can be
compared with their observed values. This is the conceptual framework
underlying the more formal versions of factor analysis \citep{cfa69} and
multidimensional scaling (MDS) \citep{borg2005}. (Note that, for this
thinking the full \pD{} data needs to be available, not just the
interpoint distances.)

@@

With our goal being to make a representation of this 2D layout that can
be lifted into high-dimensional space, the layout needs to be augmented
to include neighbour information. A simple approach would be to
triangulate the points and add edges. A more stable approach is to first
hexagonally bin the data, and connect the bin centroids. This process
serves reduce some noise in the resulting surface shown in \pD{}.

Define a hexagonal grid across the \kD{} space, using hexagons with
fixed height (\(h_b\)), width (\(w_b\)). The starting coordinates
\((x_s, y_s)\) mark the lower left of the grid. Each of the \kD{} points
will belong to a hexagonal bin. That is, for each
\(y \in \mathbfit{Y}\), we can (uniquely) identify the hexagon that the
point belongs to. Let \(h : \mathbb{R}^k \rightarrow \mathbb{R}^k\) be
the function that maps a point in \kD{} space to it's hexagonal centroid
(\(C^{(k)}\)).

@@

\subsection{\texorpdfstring{Displaying the model in
\pD{}}{Displaying the model in }}\label{displaying-the-model-in}

The final step of the algorithm is \emph{lifting} the \kD{} model back
to \pD{}, completing the \emph{model-in-the-data-space} picture. For
this, define the set \(H^k\) to be the set of points that belong to
centroid \(k\). That is, \(H^k = \{ y \; | \; g(y) = (h_x^k, h_y^k)\}\).
We use the \pD{} Euclidean mean of the points in \(H^k\) to map the
centroid \((h_x^k, h_y^k)\) to a point in \pD{}. Let the \(i\)-th
component of the \pD{} mean be

@@

with corresponding \pD{} vector \(\vec{f_i}\). Then, \(f(h\circ g)(x)\)
maps a \pD{} point to the \pD{} model estimate, completing the model
cycle.
